---
title: "Exploratory analysis for COVID paper"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
# Attach packages
library(tidyverse)
library(car)
library(RcmdrMisc)

# Read in data
source('~/Documents/IWDA Consultancy 2/data/Dimension construction/Sols_dimensions.R')
setwd("~/Documents/IWDA Consultancy 2/Report chapters")
source('summaryfuncs.R')
setwd("~/Documents/IWDA Consultancy 2/Thematic papers")

data$Age_bracket_ind <- cut(data$Age_ind, breaks = c(0, 18, 24, 34, 44, 54, 64, 74, 84, 150),
                            ordered_result = TRUE, labels = c('<18', '18-24', '25-34', '35-44', 
                                                              '45-54', '55-64', '65-74', '75-84', 
                                                              '85+'))

data$Province_ind <- car::recode(data$Province_ind, "1 = 'Central'; 2 = 'Guadalcanal'")
```

```{r, include = FALSE}
deprive <- function (raw_score){
  raw_score <- raw_score * 4
  cats_score <- ordered(cut(raw_score ,breaks=c(0,1.001,2.001,3.001,4.001),
           labels=factor(c("Most deprived","Deprived","Somewhat deprived","Least deprived")),right = FALSE))
  return(cats_score)}


# Initialise easy summary function

summarise_dim_score <- function (score, df = data, factorise = TRUE, by_province = FALSE) {
  if (factorise) {
    df$fact_score <- deprive(pull(data[score]))
  } else {
    df$fact_Score <- pull(data[score])
  }
  
  if (!by_province) {
    df %>%
    select(fact_score, Gender_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(fact_score, Gender_ind) %>%
    count() %>%
    group_by(Gender_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = fact_score, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      scale_x_discrete(limits = rev(levels(data$fact_score))) +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1))
  } else if (by_province) {
  df %>%
    select(fact_score, Gender_ind, Province_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(fact_score, Gender_ind, Province_ind) %>%
    count() %>%
    group_by(Gender_ind, Province_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = fact_score, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      scale_x_discrete(limits = rev(levels(data$fact_score))) +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1)) +
      facet_grid(cols = vars(Province_ind))
}
  }
```  


```{r}
# Initialise item score summary
summarise_item <- function(x, recode_string, by_province = FALSE) {
  if(class(recode_string) == "character") {
    data$x_recode <- car::recode(pull(data[,x]), recode_string)
  } else {
    data$x_recode <- pull(data[,x])
  }
  
  if(!by_province) {
    data %>%
    select(x_recode, Gender_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(x_recode, Gender_ind) %>%
    count() %>%
    group_by(Gender_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = x_recode, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1))
  } else if (by_province) {
    data %>%
    select(x_recode, Gender_ind, Province_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(x_recode, Gender_ind, Province_ind) %>%
    count() %>%
    group_by(Gender_ind, Province_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = x_recode, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1)) +
      facet_wrap(vars(Province_ind))
  }
}

```


## Distributions of relevant scores by gender ##

#### Sanitation ####

**Dimension score**

```{r}
summarise_dim_score('score7')
```

**Theme 1 - Toilet facilities**
```{r}
summarise_dim_score('score7.1')
```

**Theme 2 - Washing facilities**
```{r}
summarise_dim_score('score7.2')
```

**Theme 3 - Private changing facilities**
```{r}
summarise_dim_score('score7.3')
```

**A closer look at the handwashing indicator**

Here are the indicator scores for handwashing

```{r}
summarise_dim_score('score7.2.1')
```

Looking further into the component variables, most men and women have somewhere to wash their hands.

```{r}
data$Sanitation_washing_hands.place_to_wash_recode <- car::recode(data$Sanitation_washing_hands.place_to_wash, "1 = 'Yes'; 2 = 'No'; 97 = 'Refused'; else = NA")

data %>%
    select(Sanitation_washing_hands.place_to_wash_recode, Gender_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(Sanitation_washing_hands.place_to_wash_recode, Gender_ind) %>%
    count() %>%
    group_by(Gender_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = Sanitation_washing_hands.place_to_wash_recode, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1))
```

Most men and women have access to somewhere to wash there hands put there is a gender divide.

On sufficient water

```{r}
# summarise_item('Sanitation_washing_hands.enough_water', "1 = 'Yes'; 2 = 'No'; 3 = 'Sometimes'; 97 = 'Refused to answer'; else = NA")
```

Most people have sufficient water to wash their hands, but women less so than men. Is this because women need more water to wash because they're doing more domestic work that requires washing? Is it because they have less access to water?

On substances used to wash hands

```{r}
summarise_item('Sanitation_washing_hands.other_substances', "1 = 'Soap'; 2 = 'Soap substitute'; 3 = 'Just water'; 97 = 'Refused to answer'; else = NA")
```

Most people wash with soap.

#### Water ####

**Dimension score**

```{r}
summarise_dim_score('score2')
```

**Theme - Drinking water**

```{r}
summarise_dim_score('score2.1')
```

**Theme - Domestic water**

```{r}
summarise_dim_score('score2.2')
```

**Theme - Water collection threats**

```{r}
summarise_dim_score('score2.3')
```

#### Health ####

Due to issues with the prenatal health indicator, we don't have full Theme 2 data.

**Health status**

```{r}
summarise_dim_score('score4.1')
```

**Physical health status**

```{r}
summarise_dim_score('score4.1.1')
```

**Mental health status**

```{r}
summarise_dim_score('score4.1.2')
```

**Item level - Do you experience any health problems, such as headaches, dizziness, or difficulty in breathing from exposure to the smoke and fumes from your cooking fuel?**

```{r}
summarise_item('Health_status_physical.smoke', "1 = 'Yes'; 2 = 'No'; 97 = 'Refused to answer'; else = NA")
```

Interestingly, there isn't a clear pattern that shows women having more problems from cooking fuel than men.

**Theme 2 - Indicator 1 - General healthcare**

```{r}
summarise_dim_score('score4.2.1')
```

**What percentage of people needed to access healthcare but couldn't?**

```{r}
data$Explore_health_reason_not <- car::recode(data$Health_care_general_problems_notused.reason,
                                              "1:2 = 'No need / chose not'; 3:12 = 'Barriers to access'; 95 = 'Other'; 97 = 'Refused to answer' ")

summarise_item('Explore_health_reason_not', FALSE)
```

Most people who didn't access healthcare (531 respondents or 28%) but 20% of women and 18% of men couldn't access it due to barriers. The other responses here are split between barriers and no need responses roughly in the same proportion to the correctly coded answers.

**Health care problems**

```{r}
# Recode
data_health_probs <- data %>%
    select(Health_care_general_problems.clean, Health_care_general_problems.respect, Health_care_general_problems.drugs_available, Health_care_general_problems.knowledgable, Health_care_general_problems.wait, Health_care_general_problems.location, Health_care_general_problems.communication) %>%
  rename(cleaning = Health_care_general_problems.clean,
         respect = Health_care_general_problems.respect,
         drug_availability = Health_care_general_problems.drugs_available,
         knowledge = Health_care_general_problems.knowledgable,
         wait_time = Health_care_general_problems.wait,
         location = Health_care_general_problems.location,
         communication = Health_care_general_problems.communication
         ) %>%
  map_df(function (x) {
    car::recode(x, "1 = 'Yes'; 2 = 'No'; else = NA")
  })

data_health_probs$Gender_ind <- data$Gender_ind
```

```{r}
data_health_probs %>%
    gather('Problem', 'Experience', -Gender_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(Problem, Gender_ind, Experience) %>%
    count() %>%
    group_by(Gender_ind, Problem) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = Experience, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1)) +
  facet_wrap(vars(Problem))
```

I've cleaned all refused to answers (and drugs not prescribed in drug availability variable) for space reasons so percentages won't sum to zero.

We see here lots of people ran into issues with the healthcare system, this might worry us thinking about the harm caused by the COVID-19 pandemic.

## Analysis by province ##

## Distributions of relevant scores by gender ##

#### Sanitation ####

**Dimension score**

```{r}
summarise_dim_score('score7', by_province = TRUE)
```

**Theme 1 - Toilet facilities**
```{r}
summarise_dim_score('score7.1', by_province = TRUE)
```

**Theme 2 - Washing facilities**
```{r}
summarise_dim_score('score7.2', by_province = TRUE)
```

**Theme 3 - Private changing facilities**
```{r}
summarise_dim_score('score7.3', by_province = TRUE)
```

**A closer look at the handwashing indicator**

Here are the indicator scores for handwashing

```{r}
summarise_dim_score('score7.2.1', by_province = TRUE)
```

Looking further into the component variables, most men and women have somewhere to wash their hands.

```{r}
data$Sanitation_washing_hands.place_to_wash_recode <- car::recode(data$Sanitation_washing_hands.place_to_wash, "1 = 'Yes'; 2 = 'No'; 97 = 'Refused'; else = NA")

data %>%
    select(Sanitation_washing_hands.place_to_wash_recode, Gender_ind, Province_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(Sanitation_washing_hands.place_to_wash_recode, Gender_ind, Province_ind) %>%
    count() %>%
    group_by(Gender_ind, Province_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = Sanitation_washing_hands.place_to_wash_recode, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1)) +
  facet_wrap(vars(Province_ind))
```

Most men and women have access to somewhere to wash there hands put there is a gender divide.

On sufficient water

```{r}
summarise_item('Sanitation_washing_hands.enough_water', "1 = 'Yes'; 2 = 'No'; 3 = 'Sometimes'; 97 = 'Refused to answer'; else = NA", by_province = TRUE)
```

Most people have sufficient water to wash their hands, but women less so than men. Is this because women need more water to wash because they're doing more domestic work that requires washing? Is it because they have less access to water?

On substances used to wash hands

```{r}
summarise_item('Sanitation_washing_hands.other_substances', "1 = 'Soap'; 2 = 'Soap substitute'; 3 = 'Just water'; 97 = 'Refused to answer'; else = NA", by_province = TRUE)
```

Most people wash with soap.

#### Water ####

**Dimension score**

```{r}
summarise_dim_score('score2', by_province = TRUE)
```

**Theme - Drinking water**

```{r}
summarise_dim_score('score2.1', by_province = TRUE)
```

**Theme - Domestic water**

```{r}
summarise_dim_score('score2.2', by_province = TRUE)
```

**Theme - Water collection threats**

```{r}
summarise_dim_score('score2.3', by_province = TRUE)
```

#### Health ####

Due to issues with the prenatal health indicator, we don't have full Theme 2 data.

**Health status**

```{r}
summarise_dim_score('score4.1', by_province = TRUE)
```

**Physical health status**

```{r}
summarise_dim_score('score4.1.1', by_province = TRUE)
```

**Mental health status**

```{r}
summarise_dim_score('score4.1.2', by_province = TRUE)
```

**Item level - Do you experience any health problems, such as headaches, dizziness, or difficulty in breathing from exposure to the smoke and fumes from your cooking fuel?**

```{r}
summarise_item('Health_status_physical.smoke', "1 = 'Yes'; 2 = 'No'; 97 = 'Refused to answer'; else = NA", by_province = TRUE)
```

Interestingly, there isn't a clear pattern that shows women having more problems from cooking fuel than men.

**Theme 2 - Indicator 1 - General healthcare**

```{r}
summarise_dim_score('score4.2.1', by_province = TRUE)
```

**What percentage of people needed to access healthcare but couldn't?**

```{r}
data$Explore_health_reason_not <- car::recode(data$Health_care_general_problems_notused.reason,
                                              "1:2 = 'No need / chose not'; 3:12 = 'Barriers to access'; 95 = 'Other'; 97 = 'Refused to answer' ")

summarise_item('Explore_health_reason_not', FALSE, by_province = TRUE)
```

Most people who didn't access healthcare (531 respondents or 28%) but 20% of women and 18% of men couldn't access it due to barriers. The other responses here are split between barriers and no need responses roughly in the same proportion to the correctly coded answers.

**Health care problems**

```{r}
# Recode
data_health_probs <- data %>%
    select(Health_care_general_problems.clean, Health_care_general_problems.respect, Health_care_general_problems.drugs_available, Health_care_general_problems.knowledgable, Health_care_general_problems.wait, Health_care_general_problems.location, Health_care_general_problems.communication) %>%
  rename(cleaning = Health_care_general_problems.clean,
         respect = Health_care_general_problems.respect,
         drug_availability = Health_care_general_problems.drugs_available,
         knowledge = Health_care_general_problems.knowledgable,
         wait_time = Health_care_general_problems.wait,
         location = Health_care_general_problems.location,
         communication = Health_care_general_problems.communication
         ) %>%
  map_df(function (x) {
    car::recode(x, "1 = 'Yes'; 2 = 'No'; else = NA")
  })

data_health_probs$Gender_ind <- data$Gender_ind
data_health_probs$Province_ind <- data$Province_ind
```

```{r}
data_health_probs %>%
    gather('Problem', 'Experience', -Gender_ind, -Province_ind) %>%
    filter(Gender_ind =='Male' | Gender_ind == 'Female') %>%
    na.omit() %>%
    group_by(Problem, Gender_ind, Experience, Province_ind) %>%
    count() %>%
    group_by(Gender_ind, Problem, Province_ind) %>%
    mutate(prop = prop.table(n)) %>%
    ggplot(aes(x = Experience, y = prop, fill = Gender_ind)) +
      geom_col(position = 'dodge') +
      geom_text(size = 3, position = position_dodge(width = 1), aes(label = scales::percent(prop, 2), vjust = -1)) +
  facet_grid(cols = vars(Problem), rows = vars(Province_ind))
```

I've cleaned all refused to answers (and drugs not prescribed in drug availability variable) for space reasons so percentages won't sum to zero.

We see here lots of people ran into issues with the healthcare system, this might worry us thinking about the harm caused by the COVID-19 pandemic.

## Additional analysis ##

**School attainment**

```{r, include = FALSE}
data$Education_level_completion.highest_recode <- factor(data$Education_level_completion.highest, ordered = FALSE, labels = c('No schooling', 'Some primary', 'Primary', 'Some secondary/high school', 'Secondary/high school', 'Some tertiary: Vocational/ technical', 'Some tertiary: university', 'Tertiary: Vocational/technical', 'Tertiary: university', 'Refused to answer'))

data %>%
  select(Education_level_completion.highest_recode, Gender_ind, Age_bracket_ind) %>%
      filter(Gender_ind == 'Male' | Gender_ind == 'Female') %>%
      group_by(Education_level_completion.highest_recode, Gender_ind, Age_bracket_ind) %>%
      count() %>%
      group_by(Gender_ind, Age_bracket_ind) %>%
      mutate(prop = prop.table(n))%>%
      ggplot(aes(x = Age_bracket_ind, y = prop, fill = Education_level_completion.highest_recode)) +
      geom_col(position = 'stack') +
      geom_text(size = 3, position = position_stack(), aes(label = scales::percent(prop, 2), vjust = 1)) +
        facet_grid(rows = vars(Gender_ind))
```

```{r}
colPercents(xtabs(~data$Education_level_completion.highest_recode + data$Gender_ind))
```

```{r}
men <- data %>% filter(Gender_ind == 'Male')
women <- data %>% filter(Gender_ind == 'Female')
colPercents(xtabs(~men$Education_level_completion.highest_recode + men$Age_bracket_ind))
colPercents(xtabs(~women$Education_level_completion.highest_recode + women$Age_bracket_ind))

```

```{r}
colPercents(xtabs(~data$Education_level_completion.highest_recode + data$Province_ind))
```

#### Crowding ####

```{r}
data$Shelter_habitability_crowded.too_crowded_recode <- car::recode(data$Shelter_habitability_crowded.too_crowded, "1 = 'Yes'; 2 = 'No'; else = NA")

colPercents(xtabs(~data$Shelter_habitability_crowded.too_crowded_recode + data$Gender_ind))
```

```{r}
colPercents(xtabs(~data$Shelter_habitability_crowded.too_crowded_recode + data$Gender_ind + data$Province_ind))
```

